Test functions on Alabama data

### 2016-09-28 Code/Test life expectancy fxs on Alabama.Rmd
changes to LE example alabama

```{r initialize}
library(dplyr)
library(ggplot2)
library(plotly)
load("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/alabama_only.Rdata")
#example data
```

@@

```{r alabama_example}
alabama.by.strata <- 
  dat.clean.alabama %>% group_by(Year2, RaceSex, Age2) %>%
  arrange(RaceSex, Year2, Age2) %>%
  summarise(total_deaths5 = sum(Count_md5),
            Population = first(Population), 
            Race2 = first(Race2), 
            Sex2 = first(Sex2), 
            Year = first(Year),
            Age = first(Age), Age3 = first(Age3),
            StateYearRaceSex = first(StateYearRaceSex),
            nx = first(nx)) %>%
  mutate(mortality_rate = total_deaths5/Population)

first.strata <- alabama.by.strata[alabama.by.strata$StateYearRaceSex == "Alabama.1969.White.Male", ]

second.strata <- alabama.by.strata[alabama.by.strata$StateYearRaceSex == "Alabama.1969.Black.Male", ]

alabama.by.strata.cod <- dat.clean.alabama %>% group_by(Year2, RaceSex, Age2) %>%
                                               arrange(RaceSex, Year2, Age2) %>%
                                               mutate(total_deaths5 = sum(Count_md5),
                                                      cod_prop_death5 = Count_md5/total_deaths5)

first.strata.cod <- alabama.by.strata.cod[alabama.by.strata.cod$StateYearRaceSex == "Alabama.1969.White.Male", ]
second.strata.cod <- alabama.by.strata.cod[alabama.by.strata.cod$StateYearRaceSex == "Alabama.1969.Black.Male", ]

cod.tab.alabama.male.1969 <- first.strata.cod %>% ungroup() %>%
                                                  select(Age2, COD2, cod_prop_death5) %>%
                                                  rename(prop2 = cod_prop_death5)

ggplot(cod.tab.alabama.male.1969, aes(x=COD2, y=prop1)) + 
  geom_point(aes(col="White")) + geom_point(aes(y=prop2, col="Black")) +
  scale_y_continuous(name = "Proportion dying from specified cause") +
  facet_wrap(~Age2) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

system.time(life.table.whites <- data.frame(life.table(data = first.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population")))

ggplot(life.table.blacks, aes(x = Age2, y = q_x)) + geom_point() + geom_point(aes(y = R_x), col="red")

life.table.whites[, c("Age2", "total_deaths5", "Population", "R_x", "q_x", "l_x", "d_x", "L_x", "T_x", "e_x")]

system.time(
for(stratum.id in levels(droplevels(alabama.by.strata$StateYearRaceSex))) {
  assign(paste0("lt.", as.character(stratum.id)), 
         life.table(data = subset(alabama.by.strata, StateYearRaceSex == stratum.id), 
                                  age.groups = "Age2", num.ages.in.group = "nx", 
                                  death.counts = "total_deaths5", population.counts = "Population"))
}
)

>>>>

```{r initialize}
library(dplyr)
library(ggplot2)
library(plotly)
load("/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/alabama_only.Rdata")
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")
```

```{r alabama_example}
alabama.by.strata <- dat.clean.alabama %>% 
  filter(COD2 == "Cardiovascular") %>% #to only get one row per year, race, sex, age strata
  select(total_deaths5, Population, Race2, Sex2, Year, Age, StateYearRaceSex, nx) %>%
  mutate(mortality_rate = total_deaths5/Population)

first.strata <- alabama.by.strata[alabama.by.strata$StateYearRaceSex == "Alabama.1969.White.Male", ]
second.strata <- alabama.by.strata[alabama.by.strata$StateYearRaceSex == "Alabama.1969.Black.Male", ]

first.strata.cod <- dat.clean.alabama[dat.clean.alabama$StateYearRaceSex == "Alabama.1969.White.Male", ]
second.strata.cod <- dat.clean.alabama[dat.clean.alabama$StateYearRaceSex == "Alabama.1969.Black.Male", ]

#this is a table needed for the arriaga cod decomp function
cod.tab.alabama.male.1969 <- first.strata.cod %>% 
                             ungroup() %>%
                             select(Age2, COD2, cod_prop_death5) %>%
                             rename(prop2 = cod_prop_death5)

ggplot(cod.tab.alabama.male.1969, aes(x=COD2, y=prop1)) + 
  geom_point(aes(col="White")) + geom_point(aes(y=prop2, col="Black")) +
  scale_y_continuous(name = "Proportion dying from specified cause") +
  facet_wrap(~Age2) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
life.table.whites <- data.frame(life.table(data = first.strata, age.groups = "Age2", num.ages.in.group = "nx", death.counts = "total_deaths5", population.counts = "Population"))

ggplot(life.table.blacks, aes(x = Age3, y = q_x, col = "q_x: Probability")) + geom_point() + geom_point(aes(y = R_x, col="R_x: Mortality rate")) + scale_y_continuous(name = "Value") + scale_x_continuous(name = "Age") + ggtitle("Examining the relationship between the probability of dying and the mortality rate")

@@
